---
title: "Bellabeat Capstone Case Study"
author: "Adrianne Padua"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: show
execute:
  freeze: auto
---

# Introduction

This capstone project analyzes smart device usage data to explore health behaviors and trends among Bellabeat users.

Bellabeat is a high-tech company that manufactures health-focused smart products, specifically designed for women. The goal is to derive actionable insights that support Bellabeat's marketing strategies to further enhance their models, such as the **Leaf** series and **IVY+**.

# About the Data

The dataset used in this case study is the **FitBit Fitness Tracker Data** (2016), publicly available on [Kaggle](https://www.kaggle.com/datasets/arashnic/fitbit).

It includes 30 participants and provides daily activity, sleep tracking, and weight logs. Though limited in sample size and collection year, it is useful for exploratory analysis.

```{r}
#| label: load-libraries
#| message: false
library(tidyverse)
library(lubridate)

daily_activity <- read_csv("data/cleaned/dailyActivity_all.csv")
sleep_data    <- read_csv("data/cleaned/sleep_all.csv")
weight_data   <- read_csv("data/cleaned/weight_all.csv")

glimpse(daily_activity)
```

# Data Cleaning

The following steps were applied during data cleaning:

- Checked for missing or null values
- Converted date columns into proper `Date` format
- Removed duplicates
- Filtered users with complete activity and sleep logs
- Merged relevant tables for deeper insights

```{r}
# Convert ActivityDate only if not already a Date
if (!inherits(daily_activity$ActivityDate, "Date")) {
  daily_activity <- daily_activity %>%
    mutate(ActivityDate = as.Date(ActivityDate, format = "%Y-%m-%d"))
}
```

# Analysis

## Daily Activity Trends

```{r}
daily_summary <- daily_activity %>%
  summarise(
    avg_steps = mean(TotalSteps, na.rm=TRUE),
    avg_calories = mean(Calories, na.rm=TRUE)
  )
daily_summary
```

**Key Observations:**

- The average user logged fewer than the recommended **10,000 steps per day**.
- Sedentary minutes were high across most users, suggesting long inactive periods.
- Calories burned correlated with step count and active minutes.

## Sleep Behavior

```{r}
sleep_summary <- sleep_data %>%
  summarise(
    avg_minutes_asleep = mean(TotalMinutesAsleep, na.rm=TRUE),
    avg_time_in_bed = mean(TotalTimeInBed, na.rm=TRUE)
  )
sleep_summary
```

**Key Observations:**

- Many users were in bed longer than they were asleep, indicating **low sleep efficiency**.
- The average user slept **about 6.9 hours**, below the **CDC’s recommended 7+ hours** for adults.

## Weight and Activity

- Fewer participants logged weight data, limiting analysis.
- Those who did often showed irregular tracking behavior.
- There's potential to **encourage habit-building** for long-term fitness goals.

# Insights

- Users are not consistently meeting physical activity recommendations.
- Sedentary behavior dominates most user profiles.
- Sleep patterns are inconsistent, and sleep tracking could benefit from deeper analysis.
- Data logging (especially weight) is inconsistent, which limits feedback effectiveness.

# Recommendations

1. **Nutrition & Sugar Intake Logging**  
   Expand the Bellabeat app to allow **manual logging of nutrition and sugar intake**, helping users monitor how their diet affects their health goals.

2. **Integration with Glucose Meters**  
   Add support for third-party glucose monitoring devices (e.g., **OneTouch Verio**), inspired by existing solutions like **Kaiser Permanente’s KP Health Ally app**. This feature can help diabetic users (like myself) track patterns more holistically.

3. **Expanded Pregnancy & Postpartum Support for Ivy+**  
   Enhance the **IVY+** smart wearable with features tailored to **pregnant and postpartum users**, such as gentle movement goals, hydration reminders, and sleep tracking optimized for these stages of life.

4. **Condition-Specific Product Innovations**  
   Consider designing **specialized Leaf devices** (e.g., *Leaf Diabetic*) aimed at chronic conditions. These could be developed in collaboration with healthcare providers and even prescribed as part of digital wellness programs.

5. **Data-Driven Behavior Nudges**  
   Incorporate personalized **AI-based nudges** in the Bellabeat ecosystem to encourage positive habits—like hitting step goals, improving sleep hygiene, or managing sugar intake.

# Conclusion

While the dataset had limitations (small sample, 2016 data), the findings provide useful direction for Bellabeat’s product and marketing teams.  

By focusing on improving **sleep feedback**, **daily movement**, and **user engagement**, Bellabeat can better tailor its offerings to support women’s health and wellness habits through its smart products.